Improved Bounds for Online Balanced Repartitioning

### 2020-04-30 paper/disc20clustering.tex
minor
k=2 proof, clean up

We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to OPT for the first time.
Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$.
%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
paying the price of only one migration per pair.
This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
even when OPT actually pays more than that.
E.g.,
OPT can collocate two costly pairs with a single swap if
it has both pairs on the cut of  the same two clusters  (see Figure \ref{fig:TBD}).
Later, Alg collocates those pairs with two swaps.
Hence,
whenever OPT collocates a pair,
we charge OPT only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ on Alg).

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	Alg collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before Alg resets the counter.
	I.e.,
	$ C^t_{\set{u,v}} = \lambda$.
%	In case the first collocation of $\set{u,v}$ occurs at $t$,
%	if the pair is collocated in the initial configuration then
%	we set $t''=0$,
%	else,
%	we set $ t'' = -\infty, t'=0$.
	We denote the set of all requests to $\set{x,y}$
	during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$.
	If $t$ is the first time Alg collocates $\{u,v\}$ then we can charge
	OPT $\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	 or $\alpha$ for collocating the pair and serving them with no additional cost.
	Therefore in this case ,
	OPT pays $\min{\{ \lambda,\alpha \}}$.
	Else,
	let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
	in order to collocate some (intersecting) pair
	$\{x,y\} \neq \{u,v\}, x \in \{u,v\}$, 
	e.g.,
	$\{x,y\}=\{u,v'\}$.
%	I.e., $ C^{t'}_{\set{u,v'}} = \lambda$.
	Similarly,
	if $t'$ is not the first time that Alg has $\{u,v'\}$ collocated
	then let $t''<t'$ be the last time before $t'$ when Alg separates $\set{u,v'}$.
	Else,
	$\{u,v'\}$ is not collocated in the initial configuration and let $t''=0$.
	Then by these definitions,
	we have
	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
	The overall cost of requests in $\sigma_{\set{u,v}}(t',t]$,
	and the moving cost,
	incurred by Alg is
	$\lambda + 2\alpha$.
	We charge OPT the cost of serving requests in
	$\sigma_{\{u,v'\}}[t'',t']$ ($\sigma_{\{u,v\}}(t',t]$ ) 
	that it pays only within the interval $(t'', t']$ ($(t', t]$).
	Observe that for requests in $r_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$,
	OPT must pay at least
	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$
	within $(0,m]$,
	regardless of its choices.
	Based on this observation,
	we distinguish three cases of OPT's choices
	while it serves requests in	$r_t $.
	We denote its cost by
	$\mathit{Cost}_{\mathit{OPT}} (r_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
	\begin{enumerate}[label=\roman*.]
		\item
		OPT collocates $\{u,v'\}$ and $\{u,v\}$  at times  $t''$ and $t'$ respectively
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $r_t$ with no additional cost and
		$\mathit{Cost}_{\mathit{OPT}} (r_t) = 2\alpha$.
		\item
		OPT has $\{u,v\}$ ($\{u,v'\}$) already collocated at $t'$ ($t''$) as well as during $(t',t]$ ($(t'',t]$),
		and has $\{u,v'\}$ ($\{u,v\}$) separated during $(t'',t']$ ($(t',t]$).
		Hence,
		it pays only $\lambda$ for serving all requests in $r_t$ and
		$\mathit{Cost}_{\mathit{OPT}} (r_t) = \lambda$
		\item
		OPT has $\{u,v'\}$ ($\{u,v\}$) already  collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$,
		 and it collocates $\{u,v\}$ ($\{u,v'\}$) at $t'+1$ ($t''+1$) paying $\alpha$.
%		Therefore it pays only $\alpha$ during $[t'',t]$,
%		for collocating $\{u,v'\}$
%	    and for the one request to $\{u,v\}$ at $t'$.
%		I.e., it pays zero for every other  request in $r_t$.
Then,
$\mathit{Cost}_{\mathit{OPT}} (r_t) = \alpha$.
	\end{enumerate}
	 In every of these cases,
   	we have
   	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.
  	OPT may deviate from these cases,
	e.g.~by collocating a pair after partially paying for requests to this pair.
	 However,
	 any deviation only increases the cost to OPT.
	 \mahmoud{It would be better to describe and bound the "deviation" formally.}
	Therefore in any case,
 	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.

	It remains to charge both Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $m$.
	Assume $\{u,v\}$ is not collocated at $m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means Alg pays $C^{m}_{ \{u,v\} }$
	for requests to $\{u,v\}$ not accounted previously.
	If $\{u,v\}$ is not collocated in the initial configuration
	and Alg never collocates it through the sequence,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[0,m] |$.
	We charge OPT the cost
	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[0,t]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[0,t]$.
	Else,
	either $\{u,v\}$ is collocated in the initial configuration
	or Alg collocates it at some point.
	Then there exists an intersecting pair $\set{u,v'}$
	that is collocated by Alg at $t' < m$ (separating $\{u,v\}$).
	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
	Let $r_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',t]$.
	Then,
	we apply cases (i) to (iii) for $t=m$ in an analogous way,
	after replacing $\lambda$ with $C^{m}_{ \{u,v\}}$.
	Thus, 
	we obtain
	 $\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}}) 
	 \geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	 Next, we sum up all costs charged to OPT.
	 Observe that by definition,
	 $r_{t'} \cap r_t = \sigma_{\{u,v'\}}[t'',t']$.
	 This means,
	 the cost of r.h.s.~to OPT is counted at most two times in the expression
	 $\mathit{Cost}_{\mathit{OPT}} (r_{t'}) + \mathit{Cost}_{\mathit{OPT}} (r_t)$.
	 Hence,
	 for all collocations performed by Alg and for $t=m$,
	 OPT pays at least 
	 $\sum_{ 0 < t \leq |\sigma|} \mathit{Cost}_{\mathit{OPT}} (r_t) / 2$.
	Let
	$T := \{ 0 < t \leq m ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
	By setting $\lambda= \alpha$
	we obtain
\begin{align*}
	\mathit{Cost}_{\mathit{ALG}} (\sigma)	/
	 \mathit{Cost}_{\mathit{OPT}} (\sigma)
	 &=
	 \Big(
	 \sum_{ t \in T}(\lambda + 2\alpha) +
	 \sum_{\{u,v\}} C^{m}_{\{u,v\}}
	 \Big)	 \big /
	 \Big(
	 \sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (r_t) / 2
	 + \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})  / 2 
	 \Big)	\\
	 &\leq
	 \Big(
	 \sum_{ t \in T}(\lambda + 2\alpha) +
	 \sum_{\{u,v\}} C^{m}_{\{u,v\}}
	 \Big)	 \big /
	 \Big(
	 \sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  / 2
	 + \sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \}  / 2 } 
	 \Big)	\\
	 &=
	 \Big(
	\sum_{ t \in T} 3\alpha +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
	\Big)	 \big /
	\Big(
	\sum_{ t \in T} \alpha  / 2
	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  / 2 
	\Big)	\leq 6
\end{align*}.
	%	The claim follows from here.
\end{proof}

>>>>

We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to OPT for the first time.
Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$.
%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
paying the price of only one migration per pair  (see Figure \ref{fig:TBD}).
%This observation suggests that we can't charge OPT more than $\alpha$ per collocation.
%E.g.,
%OPT can collocate two costly pairs with a single swap if
%it has both pairs on the cut of  the same two clusters.
%Later, Alg collocates those pairs with two swaps.
Hence,
whenever OPT collocates a pair,
we charge it only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ on Alg).

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	Alg collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before Alg resets the counter.
	I.e.,
	$ C^t_{\set{u,v}} = \lambda$.
%	In case the first collocation of $\set{u,v}$ occurs at $t$,
%	if the pair is collocated in the initial configuration then
%	we set $t''=0$,
%	else,
%	we set $ t'' = -\infty, t'=0$.
	We denote the set of all requests to a pair $\set{x,y}$ that arrive
	during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$.
	If $t$ is the first time that Alg collocates $\{u,v\}$ then we  charge
	OPT $\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	 or $\alpha$ for collocating the pair and serving them with no additional cost.
	Therefore in this case ,
	OPT pays $\min{\{ \lambda,\alpha \}}$.
	Else,
	let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
	in order to collocate some (intersecting) pair
	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
	e.g.,
	$\{x,y\}=\{u,v'\}$.
%	I.e., $ C^{t'}_{\set{u,v'}} = \lambda$.
	Similarly,
	if $t'$ is not the first time that Alg  collocates $\{u,v'\}$ 
	then let $t''<t'$ be the last time before $t'$ when Alg separates $\set{u,v'}$.
	Else,
	$\{u,v'\}$ is not collocated in the initial configuration and we let $t''=0$.
	By these definitions,
	we have
	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
	The overall cost of requests in $\sigma_{\set{u,v}}(t',t]$,
	and the moving cost,
	incurred by Alg is
	$\lambda + 2\alpha$.
	Observe that for requests in $r_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$,
	OPT must pay a total cost of serving and (or) migration at least
	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$
	within $(0,m]$,
	regardless of its choices.
	%% 
	However,
	we cannot charge this cost to OPT repeatedly,
	i.e.,
	when Alg collocates the two pairs a second time.
	In particular,
	we charge OPT a cost incurred by requests in
	$\sigma_{\{u,v'\}}(t'',t']$ and $\sigma_{\{u,v\}}(t',t]$,
	if and only if  OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
	respectively.	
	Latter,
	this ensures us that we do not charge OPT for a cost more than twice.
	%%
	Based on this observation,
	we distinguish three cases of OPT's choices
	while it serves requests in	$r_t $.
	We denote its cost by
	$\mathit{Cost}_{\mathit{OPT}} (r_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
	\begin{enumerate}[label=\roman*.]
		\item
		OPT collocates $\{u,v'\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $r_t$ with no additional cost and
		$\mathit{Cost}_{\mathit{OPT}} (r_t) = 2\alpha$.
		\item
		OPT has $\{u,v\}$ ($\{u,v'\}$) already collocated at $t'$ ($t''$) as well as during $(t',t]$ ($(t'',t]$),
		and has $\{u,v'\}$ ($\{u,v\}$) separated during $(t'',t']$ ($(t',t]$).
		Hence,
		it pays only $\lambda$ for serving all requests in $r_t$ and
		$\mathit{Cost}_{\mathit{OPT}} (r_t) = \lambda$
		\item
		OPT has $\{u,v'\}$ ($\{u,v\}$) already  collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$,
		 and it collocates $\{u,v\}$ ($\{u,v'\}$) at $t'+1$ ($t''+1$) paying $\alpha$.
%		Therefore it pays only $\alpha$ during $[t'',t]$,
%		for collocating $\{u,v'\}$
%	    and for the one request to $\{u,v\}$ at $t'$.
%		I.e., it pays zero for every other  request in $r_t$.
Then,
$\mathit{Cost}_{\mathit{OPT}} (r_t) = \alpha$.
	 In every of these cases,
   	we have
   	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.
  	OPT may deviate from these cases,
	e.g.~by collocating a pair after partially paying for requests to this pair.
	 However,
	 no deviation can decrease the cost to OPT.
	 \mahmoud{It would be better to describe and bound the "deviation" formally.}
	Therefore in any case,
 	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.

	It remains to charge both Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means Alg pays $C^{m}_{ \{u,v\} }$
	for requests to $\{u,v\}$ so far unaccounted.
	If $\{u,v\}$ is not collocated in the initial configuration
	and Alg never collocates it through the sequence,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	We charge OPT the cost
	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[1,t]$.
	Else,
	either $\{u,v\}$ is collocated in the initial configuration
	or Alg collocates it at some point.
	Then there exists an intersecting pair $\set{u,v'}$
	that is collocated by Alg at $t' < m$ (separating $\{u,v\}$).
	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
	Let $r_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	we apply cases (i) to (iii) in an analogous way,
	 replacing $\lambda$ with $C^{m}_{ \{u,v\}}$ in the analysis.
	Thus, 
	we obtain
	 $\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}}) 
	 \geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	 Next, we sum up all costs charged to OPT.
	 Observe that by definition,
	 $r_{t'} \cap r_t = \sigma_{\{u,v'\}}[t'',t']$.
	 This means,
	 the cost of r.h.s.~to OPT is counted at most twice in the expression
	 $\mathit{Cost}_{\mathit{OPT}} (r_{t'}) + \mathit{Cost}_{\mathit{OPT}} (r_t)$.
	 Hence,
	 for all collocations performed by Alg,
	 and for final costs at $t=m$,
	 OPT pays at least 
	 $(
	 \sum_{ 0 < t \leq |\sigma|} \mathit{Cost}_{\mathit{OPT}} (r_t) +
	 \sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})
	 ) / 2$.
	Let
	$T := \{ 0 < t \leq m ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
	By setting $\lambda= \alpha$,
	we obtain
\begin{align*}
	\mathit{Cost}_{\mathit{ALG}} (\sigma)	/
	 \mathit{Cost}_{\mathit{OPT}} (\sigma)
	 &=
	 \Big(
	 \sum_{ t \in T}(\lambda + 2\alpha) +
	 \sum_{\{u,v\}} C^{m}_{\{u,v\}}
	 \Big)	 \big /
	 \frac{1}{2}
	 \Big(
	 \sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (r_t) 
	 + \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})
	 \Big)	\\
	 &\leq
	 \Big(
	 \sum_{ t \in T}(\lambda + 2\alpha) +
	 \sum_{\{u,v\}} C^{m}_{\{u,v\}}
	 \Big)	 \big /
		\frac{1}{2}
	 \Big(
	 \sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
   \sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
	 \Big)	\\
	 &=
	 2\Big(
	\sum_{ t \in T} 3\alpha +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
	\Big)	 \big /
	\Big(
	\sum_{ t \in T} \alpha  
	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
	\Big)	< 6
\end{align*}.
	%	The claim follows from here.
\end{proof}
